---
title: "Project 1"
author: "Ethan Michaels"
date: "January 31, 2024"
output: pdf_document
---
#Loading in libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
```

#Opening the CSV file and Cleaning
```{r}
data <- read.csv('franchise values.csv')
```

```{r}
format(data,big.mark=",",scientific=FALSE)
```

#Cleaning Data to remove any teams value that are above our price range
```{r}
data= data %>% filter(Value<4000000000)
```

```{r}
summary(data)
```

```{r}
cor(data[,c('Value','Revenue','Debt.Value','Income','Population')])
```
#Creating a new variable to classify teams and year in one varible to have an easier time cleaning up the data
```{r}
data$TeamYear <- paste(data$Team,data$Year)
```

#Sorting the data to show data by league by year  
```{r}
data_sorted <- data %>% arrange(desc(League),TeamYear,Year)
```

#Creating a variable for value to revenue ratio
```{r}
data_sorted$Value_Revenue = data_sorted$Value/ data_sorted$Revenue
data_sorted$Revenue_Value = data_sorted$Revenue/ data_sorted$Value
data_sorted$Debt = data_sorted$Value*data_sorted$Debt.Value
```

#Making all varaibles with big numbers to show numbers by millions
```{r}
data_sorted$Revenue <- data_sorted$Revenue/1000000
data_sorted$Value <- data_sorted$Value/1000000
data_sorted$Income <- data_sorted$Income/1000000
data_sorted$Population <- data_sorted$Population/1000000
data_sorted$Debt <- data_sorted$Debt/1000000
data_sorted$Income_Revenue <- data_sorted$Income/data_sorted$Revenue

```

```{r}
ggplot(data_sorted,aes(x=Value,y=Revenue)) +
  geom_point(aes(color=factor(League)))
```

#Taking the mean value of the Revenue_Value Ratio for each League Each Year
```{r}
data_summarise <- group_by(data_sorted,League,Year)
data_summarise <-summarise(data_summarise,avg=mean(Revenue_Value))
```

#Creating a barplot to show the difference between leagues in each years average Value to revenue ratio
```{r}
ggplot(data_summarise,aes(fill=data_summarise$League,y=data_summarise$avg,x=data_summarise$Year,xlab="Year"))+
  geom_bar(position="dodge",stat="identity") +
  labs(x="Year",y="Average Revenue to Revenue Ratio",fill="League")
```

#######################################################################################################################################################
#After comparing the data between sports I am Taking out 2020 due to lack of information, and MLS data since there is only 1 year worth of data
```{r}
data_sorted <- data_sorted[!data_sorted$Year == 2020, ]
data_sorted <- data_sorted[!data_sorted$League == 'MLS', ]
data_sorted2 <- data_sorted[!data_sorted$Year == 2021,]
```

#SplittingDataframes by League
```{r}
nfl_data <- data_sorted[data_sorted$League=='NFL',]
mlb_data <- data_sorted[data_sorted$League=='MLB',]
nba_data <- data_sorted[data_sorted$League=='NBA',]
nhl_data <- data_sorted[data_sorted$League=='NHL',]
```

```{r}
nfl_data2 <- nfl_data
na.omit(nfl_data2)
nfl_data2 <- nfl_data2[!nfl_data2$Year == 2021, ]
mlb_data2 <- mlb_data[!mlb_data$Year == 2021, ]
nhl_data2 <- nhl_data[!nhl_data$Year == 2021, ]
nba_data2 <- nba_data[!nba_data$Year == 2021, ]
```

```{r}
ggplot(data_sorted,aes(x=Value,y=Revenue)) +
  geom_point(aes(color=factor(League)))
```

```{r}
ggplot(data_sorted,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point(aes(color=factor(League)))
```

```{r}
ggplot(data_sorted,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point(aes(color=factor(League)))
```


```{r}
ggplot(nfl_data2,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

```{r}
ggplot(mlb_data2,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

```{r}
ggplot(nhl_data2,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

```{r}
ggplot(nba_data2,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

```{r}
ggplot(data_sorted2,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

```{r}
data_sorted3 <- data_sorted2[data_sorted2$Revenue_Value>data_sorted2$Debt.Value, ]
```

```{r}
data_sorted3 
```

```{r}
data_sorted4 <- data_sorted2[data_sorted2$Value <1000, ]
data_sorted4
```

```{r}
ggplot(data_sorted4,aes(x=Revenue_Value,y=Debt.Value)) +
  geom_point()+
  geom_text_repel(aes(label=Team),size=3,vjust=1)+
  theme()
```

#Checking the correlation of all our variables
```{r}
cor(data_sorted[,c('Value','Revenue','Debt.Value','Income','Population','Income','X..of.Seasons','Debt'
                   )])
```

#Creating a basic linear model for all Leagues excluding any N/As
#USE
```{r}
linear_fit <- lm(Value ~ Revenue+Income+Population+Debt, data=data_sorted2)
summary(linear_fit)

```

#NFL Regression
```{r}
linear_fit_log <- lm(Value ~ Revenue + Income + Debt + Population, data=nfl_data2)
summary(linear_fit_log)
```

#MLB_Data
```{r}
linear_fit_log <- lm(Value ~ Revenue + Income+ Debt + Population, data=mlb_data2)
summary(linear_fit_log)
```

#NBA Data
```{r}
linear_fit_log <- lm(Value ~ Revenue + Income + Population, data=nba_data2)
summary(linear_fit_log)
```

#NHL
```{r}
linear_fit_log <- lm(Value ~ Revenue + Income + Population, data=nhl_data2)
summary(linear_fit_log)
```

